---
title: "Climate Colonialism: Gather.Rmd"
author: "Winona Guo"
date: "10/14/2020"
output: html_document
---

These are my datasets:
  
1) Colonial History Dataset
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/5EMETG

2) Measuring the Impacts of Colonialism: A New Data Set for the Countries of Africa and Asia
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/UQZFYA

3) Countries Vulnerable to the Climate
https://www.cgdev.org/publication/dataset-vulnerability-climate-change

4) C02 Emissions Per Country
https://github.com/datasets/co2-fossil-by-nation

Below are resources that were helpful to reference as I made this project:

https://shiny.rstudio.com/reference/shiny/1.5.0/
https://shiny.rstudio.com/tutorial/
https://github.com/wyatthurt/gov-50-class-scripts/blob/main/class-2020-09-22/week_3_A.Rmd
https://rdrr.io/cran/maptools/man/wrld_simpl.html
https://www.davidkane.info/files/final_projects.html
https://github.com/wyatthurt/transboundary-water-conflict/tree/master/shiny_app
https://www.r-graph-gallery.com/
https://benalexkeen.com/creating-a-timeline-graphic-using-r-and-ggplot2/

Sources cited:
https://www.opendemocracy.net/en/opendemocracyuk/to-fix-climate-crisis-we-must-acknowledge-our-imperial-past/
https://psmag.com/social-justice/how-the-gnd-can-avoid-climate-colonialism

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(shiny)
library(shinythemes)
library(tidyverse)
library(janitor)

# Janitor is used to the clean the names within my data.

library(readr)
library(readxl)
library(leaflet)
library(maptools)
library(rgdal)
library(dplyr)
library(tidymodels)
library(rstanarm)
library(gt)
library(broom.mixed)
library(gganimate)
library(gtsummary)

```

```{r Loading Data}

colonialism <- read_xls("shiny_app/raw_data/colonialism.xls") %>%
  clean_names() %>%
  rename(colonizer = main_colonial_motherland_source_ziltener_k_ynzler,
         onset = onset_of_colonialism_source_ziltener_k_ynzler_2008,
         end = end_of_colonialism_source_ziltener_k_ynzler_2008) 

climaterisk <- read_xls("shiny_app/raw_data/vulnerability.xls", skip = 2) %>%
  clean_names() %>%
  rename(vulnerable = climate_vulnerability_cv_cdi_adj_for_income_regulation,
         governance = kkm_regulatory_quality_score_2008, 
         income = income_per_capita_us_ppp_2008, 
         area = area_sq_km,
         extreme_weather = climate_vulnerability_wcv_wcdi_adj_for_income_regulation,
         sea_level_rise = climate_vulnerability_scv_scdi_adj_for_income_regulation,
         agr_prod_loss = climate_vulnerability_acv_acdi_adj_for_income_regulation) %>%
  mutate(country = str_to_title(country)) %>%
  mutate(country = recode(country, 'Korea, Rep.' = 'Korea, Republic of',
                          "Korea, Dem. Rep." = "Korea, Democratic People's Republic of",
                          'Taiwan (China)' = 'Taiwan',
                          'Antigua And Barbuda' = 'Antigua and Barbuda',
                          'Bahamas, The' = 'Bahamas',
                          'Slovak Republic' = 'Slovakia',

                          # Starting here are the ones I am renaming so that it aligns with the wrldsimplmod countries' names. 

                          'Iran, Islamic Rep.' = 'Iran (Islamic Republic of)',
                          'Bosnia And Herzegovina' = 'Bosnia and Herzegovina',
                          'Myanmar' = 'Burma',
                          'Congo, Rep.' = 'Congo',
                          'Congo, Dem. Rep.' = 'Democratic Republic of the Congo',
                          'Egypt, Arab Rep.' = 'Egypt',
                          'Yemen, Rep.' = 'Yemen',
                          'French Guians' = 'French Guiana',
                          'Micronesia, Fed. Sts.' = 'Micronesia, Federated States of',
                          'Gambia, The' = 'Gambia',
                          "Cote D'ivoire" = "Cote d'Ivoire",
                          'Kyrgyz Republic' = 'Kyrgyzstan',
                          'Falkland Islands' = 'Falkland Islands (Malvinas)',
                          "Lao Pdr" = "Lao People's Democratic Republic",
                          'Libya' = 'Libyan Arab Jamahiriya',
                          'Macedonia, Fyr' = 'The former Yugoslav Republic of Macedonia',
                          'Hong Kong Sar, China' = 'Hong Kong',
                          'Isle Of Man' = 'Isle of Man',
                          'Macao Sar, China' = 'Macau',
                          'West Bank And Gaza' = 'Palestine',
                          
                          # I feel somewhat like I am simplifying decades of history into one line of code!
                          
                          'Serbia And Montenegro' = 'Serbia',
                          'Moldova' = 'Republic of Moldova',
                          'Russian Federation' = 'Russia',
                          'St. Kitts And Nevis' = 'Saint Kitts and Nevis',
                          'St. Lucia' = 'Saint Lucia',
                          'Trinidad And Tobago' = 'Trinidad and Tobago',
                          'Sao Tome And Principe' = 'Sao Tome and Principe',
                          'Tanzania' = 'United Republic of Tanzania',
                          'St. Vincent And The Grenadines' = 'Saint Vincent and the Grenadines',
                          'Venezuela, Rb' = 'Venezuela',
                          'Vietnam' = 'Viet Nam',
                          'Virgin Islands (U.s.)' = 'United States Virgin Islands',
                          'Wallis And Futuna' = 'Wallis and Futuna Islands',
                          'Pitcairn' = 'Pitcairn Islands',
                          'St. Pierre And Miquelon' = 'Saint Pierre and Miquelon',
                          'St. Helena' = 'Saint Helena',
                          'Turks And Caicos Islands' = 'Turks and Caicos Islands',
                          'Svalbard And Jan Mayen' = 'Svalbard')) %>%
    mutate(vulnerable = replace(vulnerable, 188, 42.000))

# BIG IMPORTANT NOTE: I replaced Somalia's 100.000 climate risk rate with 42.000 (near Burma's), because it is a single dramatic outlier that changes the later graph. I will be sure to note this via text inside my Shiny app!

# Mapping these data sets (and doing the corresponding recoding work) requires navigating a lot of complex history that I'm having difficulty summarizing into one number. For instance, in the countries data, we have Yemen listed as being decolonized in 1990, but Yemen People’s Republic decolonized in 1967. In the climate data, there is one one 'Yemen Rep.' Which name maps on to which? I ended up choosing to rename Yemen Rep. as Yemen, both to match the later colonization date and also to match wrldsimplmod.

countries <- read_xls("shiny_app/raw_data/countries_files/icowcol.xls") %>%
  mutate(Indep = str_sub(Indep, 1, 4)) %>%
  mutate(IndepTC = str_sub(IndepTC, 1, 4)) %>%
  mutate(Name = recode(Name, 'United States of America' = 'United States',
                       'Tunisia (postcolonial)' = 'Tunisia',
                       'St. Kitts and Nevis' = 'Saint Kitts and Nevis',
                       'St. Vincent and the Grenadines' ='Saint Vincent and the Grenadines',
                       'Prussia / Germany' = 'Germany',
                       'Fed. States of Micronesia' = 'Micronesia, Federated States of',
                       'Iran' = 'Iran (Islamic Republic of)',
                       'Brunei' = 'Brunei Darussalam',
                       'Austria-Hungary' = 'Austria',
                       "Laos" = "Lao People's Democratic Republic",
                       'Egypt (poat-colonial)' = 'Egypt',
                       'South Korea' = 'Korea, Republic of',
                       "North Korea" = "Korea, Democratic People's Republic of",
                       'Moldova' = 'Republic of Moldova',
                       'Serbia / Yugoslavia' = 'Serbia',
                       'Tanzania' = 'United Republic of Tanzania',
                       'Vietnam' = 'Viet Nam',
                       'Sardinia / Italy' = 'Italy', 
                       "St. Lucia" = "Saint Lucia",
                       'Morocco (postcolonial)' = 'Morocco',
                       'Libya' = 'Libyan Arab Jamahiriya',
                       'Estonia (post-Soviet)' = 'Estonia',
                       'Latvia (post-Soviet)' = 'Latvia',
                       'Lithuania (post-Soviet)' = 'Lithuania',
                       'Macedonia' = 'The former Yugoslav Republic of Macedonia',
                       'Myanmar' = 'Burma',
                       'Syria (post-UAR)' = 'Syrian Arab Republic',
                       'East Timor' = 'Timor-Leste')) %>%
  mutate(From = replace_na(From, "Not Specified")) %>%
  mutate(From = recode(From, "200" = "United Kingdom",
                       "2" = "United States",
                       "220" = "France",
                       "41" = "Haiti",
                       "230" = "Spain",
                       "89" = "Not Specified",
                       "100" = "Colombia",
                       "210" = "Netherlands",
                       "235" = "Portual",
                       "-9" = "Not Specified",
                       "255" = "Germany",
                       "300" = "Austria-Hungary",
                       "365" = "Russia",
                       "315" = "Czechoslovakia",
                       "345" = "Yugoslavia",
                       "640" = "Turkey",
                       "380" = "Sweden", 
                       "390" = "Denmark",
                       "432" = "Mali",
                       "590" = "Mauritius",
                       "698" = "Oman",
                       "710" = "China",
                       "820" = "Malaysia",
                       "850" = "Indonesia",
                       "900" = "Australia",
                       "920" = "New Zealand",
                       "730" = "Korea",
                       "678" = "Yemen Arab Republic",
                       "651" = "Egypt",
                       "640" = "Turkey"))

emissions <- read_csv("shiny_app/raw_data/emissions.csv", 
                      col_types = cols(
                        Year = col_double(),
                        Country = col_character(),
                        Total = col_double(),
                        `Solid Fuel` = col_double(),
                        `Liquid Fuel` = col_double(),
                        `Gas Fuel` = col_double(),
                        Cement = col_double(),
                        `Gas Flaring` = col_double(),
                        `Per Capita` = col_double(),
                        `Bunker fuels (Not in Total)` = col_double()
                      )) %>%
  group_by(Country) %>%
  mutate(sum = sum(Total)) %>%
  slice(1) %>%
  select(Country, sum) %>%
  mutate(Country = str_to_title(Country)) %>%
  mutate(Country = recode(Country, "United States Of America" = "United States",
                          "China (Mainland)" = "China",
                          "Russian Federation" = "Russia",
                          "Islamic Republic Of Iran" = "Iran (Islamic Republic of)",
                          "Plurinational State Of Bolivia" = "Bolivia",
                          "Myanmar (Formerly Burma)" = "Burma",
                          "United Republic Of Tanzania" = "United Republic of Tanzania",
                          "Libyan Arab Jamahiriyah" = "Libyan Arab Jamahiriya",
                          "French West Africa" = "Western Sahara",
                          "Cote D Ivoire" = "Cote d'Ivoire",
                          "Republic Of Cameroon" = "Cameroon",
                          "Democratic Republic Of The Congo (Formerly Zaire)" = "Democratic Republic of the Congo",
                          "United Republic Of Tanzania" = "United Republic of Tanzania",
                          "Lao People S Democratic Republic" = "Lao People's Democratic Republic",
                          "Republic Of Korea" = "Korea, Republic of", 
                          "Democratic People S Republic Of Korea" = "Korea, Democratic People's Republic of",
                          "Antarctic Fisheries" = "Antarctica",
                          "France (Including Monaco)" = "France",
                          "Italy (Including San Marino)" = "Italy")) %>%
    mutate(sum = ifelse(sum == 102510260, 60000000, sum))

# BIG IMPORTANT NOTE AGAIN: I replaced the U.S. total emissions of 102510260, a dramatic outlier, to be 60,000,000, closer to the second largest # of 47649834 (China). 

# The below joins the colonialism data with climate risk data!

fit_mod <- inner_join(countries, climaterisk, by = c("Name" = "country")) %>%
  select(-COWsys,
         -GWsys,
         -Notes) 

 
```

```{r CO2 and Climate Risk}

# Model 3: Is there a predictive relationship between CO2 emissions and climate risk?

# Trying to produce total emissions for each country between 1751 and 2014. 

model_emissions <- inner_join(emissions, climaterisk, by = c("Country" = "country")) 
 
model_emissions_plot <- model_emissions %>%
  
# Without this transformation in the next line, I always get the error of
# "NaNs"! Online I'm finding that no negative numbers are allowed in the log
# transformation... I'm also assuming 0 is not allowed since it's a log
# transformation.
  
 mutate(vulnerable = ifelse(vulnerable %in% c(0.000, -0.013, -0.16, -0.237, -0.035, -0.440, -0.016), 0.001, vulnerable)) %>%
 ggplot(aes(x = sum, y = vulnerable)) +
 geom_point(na.rm = TRUE, color = "red") +
 scale_x_log10() + 
  
  # This gives me not 1 x 10^7 but log 1 x 10^7.
  
 scale_y_log10() +
 labs(title = "Total Emissions vs. Climate Vulnerability",
       y = "Climate Vulnerability (log)",
       x = "Total Emissions 1751-2014 (log)")

model_emissions_plot

# Saving the model...

model_emissions_stan <- stan_glm(formula = vulnerable ~ sum,
                                  data = model_emissions, 
                                  refresh = 0, 
                                  seed = 8)

saveRDS(model_emissions_stan, file = "shiny_app/model_emissions_stan.rds")

# There is no substantive predictive effects here!

```

```{r Defining Colonized}

# What countries count as colonized? The data set could be split many ways... I will run the RMSE of each in my model to determine which is best. This information will go on later Shiny tabs, but I am including it as the first part of this .RMD so my data sets are created and defined!

# First option: all those coded as decolonial countries in the data. This
# includes countries like the United States, credited as beginning a Western
# history of decolonization starting in the 1770s. This is based on a
# generalized public history: https://en.wikipedia.org/wiki/Decolonization

all_decolonized <- fit_mod %>%
 mutate(status = ifelse(Type == 2, "Yes", "No")) %>%
 select(Name, Indep, Type, climate_drivers_cdi, vulnerable, income, governance, area, world_region, world_sub_region, status)

# Second option: Revising the data set with historical particularities.
# Including all countries marked by secession or partition, like Guatemala and
# North/South Korea. Many of these countries indeed also have an imperial
# history, though not all in the same or similar ways. Also, removing the U.S.
# because of its later emergence as an imperial power, and Canada for
# participating in a related project. None of the other countries on the list
# are comparable to their wealth and power (especially of the U.S.) in recent
# centuries.

decolonial_revised <- fit_mod %>%
  mutate(status = ifelse(Type %in% c(2, 3, 4), "Yes", "No")) %>%
  filter(Name != "United States") %>%
  filter(Name != "Canada")

# Third option: Revised data set above, but also adding the condition of
# countries that were decolonized after 1850, when industrial-era warming
# accelerated. I hypothesize this may better capture the predictive effect of
# colonization on climate, since colonialism is ongoing in a period of
# increasing climate change. https://www.nature.com/articles/nature19082

nineteenth_century <- decolonial_revised %>%
  filter(Indep >= 1850)


```

```{r Colonialism + Climate Graphs}

# PLOTTING AVERAGE CLIMATE RISK BY COLONIAL STATUS

status_plot <- all_decolonized %>%
 group_by(status) %>%
 mutate(avg_risk = mean(vulnerable)) %>% 
  
# Above I am calculating a very generalized average climate risk for countries marked as colonized vs. not. 
  
 select(Name, status, world_region, avg_risk) %>%
 slice(1) %>%
 ggplot(mapping = aes(x = status, y = avg_risk)) + 
 geom_col(fill = "darkseagreen") +
 labs(title = "Climate Risk for Independent vs. Colonized Countries", 
      x = "Colonial Status",
      y = "Avg. Climate Vulnerability",
      subtitle = "As of 2011",
      caption = "The average climate risk for independent countries is 1.53, 
      and the average for colonized countries is 4.36.") +
 theme_classic() 

status_plot

# BOXPLOT

box_plot <- all_decolonized %>%
 group_by(status) %>%
  
# Above I am calculating a very generalized average climate risk for countries marked as colonized vs. not. 

 ggplot(mapping = aes(x = status, y = vulnerable)) + 
 geom_boxplot(fill = "darkseagreen", outlier.shape = NA) +
 coord_cartesian(ylim = c(0, 11)) +
 labs(title = "Climate Risk for Independent vs. Colonized Countries", 
      subtitle = "Removing outliers",
      x = "Colonial Status",
      y = "Avg. Climate Vulnerability") +
 theme_classic()

box_plot

# PLOTTING BY WORLD REGION

world_region_plot <- all_decolonized %>%
 group_by(world_region) %>%
 mutate(avg_risk = mean(vulnerable)) %>% 
  
# Above I am calculating averages by continent instead of by colonial status.
  
 select(Name, status, world_region, avg_risk) %>%
 slice(1) %>%
 ggplot(mapping = aes(x = reorder(world_region, avg_risk), y = avg_risk)) + 
 geom_col(fill = "orange") +
 labs(title = "Climate Risk Across Six Continents", 
      x = "Continents",
      y = "Average Climate Vulnerability") +
 theme_classic() 

world_region_plot


```

```{r Models}

# Model 1: Is there a predictive relationship between colonialism and climate risk today?

model_1 <- stan_glm(formula = vulnerable ~ status,
                data = all_decolonized, 
                refresh = 0, 
                seed = 8)

# Results: Estimate   SD
# Intercept	1.563682	0.797341		
# statusYes	2.815612	1.004868	

model_2 <- stan_glm(formula = vulnerable ~ status,
                data = decolonial_revised, 
                refresh = 0, 
                seed = 8)

# (Intercept)	2.097084	1.317358		
# statusYes	1.463461	1.380297	


model_3 <- stan_glm(formula = vulnerable ~ status,
                data = nineteenth_century, 
                refresh = 0, 
                seed = 8) 

# (Intercept)	2.589422	2.533934		
# statusYes	 1.455264	 2.646424

# Choosing Model 1 because it has the lowest SD. 

# Graphing it below.

pp <- model_1 %>%
            as_tibble() %>%
            rename(mu = `(Intercept)`, 
                   status = "statusYes") %>%
            ggplot(aes(x = mu)) +
            geom_histogram(aes(y = after_stat(count/sum(count))), 
                           bins = 100) +
            labs(title = "Posterior Probability Distribution",
                 subtitle = "Average climate vulnerability among countries in 2011",
                 x = "Climate Vulnerability",
                 y = "Probability") +
            theme_classic()
          
          pp

# Model 2: Adding the variable of world region is interesting...

model_world <- stan_glm(formula = vulnerable ~ status + world_sub_region,
                data = all_decolonized, 
                refresh = 0, 
                iter = 2500, 
                
                # Without the above line, I get an error message that my Bulk
                # ESS (effective sample size) may be too low. The suggested
                # solution was found here:
                # https://mc-stan.org/misc/warnings.html#bulk-ess
                
                seed = 8)

saveRDS(model_world, file = "shiny_app/model_world.rds")


# So is governance.

model_gov <- stan_glm(formula = vulnerable ~ status + governance,
                data = all_decolonized, 
                refresh = 0, 
                seed = 8) 

saveRDS(model_gov, file = "shiny_app/model_gov.rds")


```

* Mathemtical representation for the first model. I am considering risk as a function of colonial status. How much does colonialism affect risk?

$$ reg_{risk} = \beta_0 + \beta_1x_{status, i} + \epsilon $$
```{r Tab 4 Leaflet}

# Modifying Polygons: world countries from the maptools package.

data(wrld_simpl)

wrldsimplmod <- wrld_simpl

# Merging with my data on climate risk. Unfortunately some countries are left with NA's because there is no climate risk data. An example: the South Georgia South Sandwich Islands!

wrldsimplmod@data <-
  
   left_join(wrld_simpl@data, climaterisk, by = c("NAME" = "country"))

saveRDS(wrldsimplmod, "shiny_app/joined.rds")

# Making for the second graph.

secondmod <- wrld_simpl

secondmod@data <-
  
   left_join(wrld_simpl@data, emissions, by = c("NAME" = "Country"))

saveRDS(secondmod, "shiny_app/joined2.rds")

```

``` {r Unused Code and Future Ideas}


# graph onset and end of colonialism for different countries
# graph the 4 variables change over that period
# graph france vs. uk whether they have different effects on climate risk
# countries on left, geomline line segments on y from onset/end, color coded by motherland
# PLOT: COLONIALISM & INCOME

# PLOT 2: NUMBER OF YEARS COLONIZED (African and Asian Countries)

#  q1 <- colonialism %>%
#  select(country_name, colyears) 

#  q2 <- climaterisk %>%
#  select(country, vulnerable, world_sub_region) %>%
#  subset(country != "Somalia") %>%
#  subset(country != "Burundi")

#  I removed these two outliers because they were so extreme that the rest of the graph was difficult to decipher.

#  q3 <- inner_join(q1, q2, by = c("country_name" = "country")) 

#  colonize <- q3 %>%
#  ggplot(aes(x = colyears, y = vulnerable, color = world_sub_region)) +
#  geom_point(alpha = 0.7) +
#  labs(title = "Countries by Years Colonized and Vulnerability to Climate Risk",
#       x = "Years Colonized",
#       y = "Vulnerability to Climate Risk",
#       caption = "This graph indicates there is not a significant correlation
#       overall between the years a country has been colonized, 
#       and its current vulnerability to climate risk.") +
#  theme_bw() + 
#  scale_color_discrete("World Region")

#  PLOT: TYPE OF INDEPENDENCE (x) ON CLIMATE RISK (y)
      
#    types <- fit_mod %>%
#     ggplot(aes(x = Type, y = vulnerable)) +
#     geom_col(fill = "orange") +
#     labs(title = "The Effect of Different Types of Colonialism on Climate Risk",
#           y = "Climate Risk",
#           subtitle = "Key: 1 = Formation, 2 = Decolonization, 3 = Secession, 4 = Partition",
#           caption = "This graph shows that decolonialization and occuption by a foreign power 
#           has a significantly greater effect on risk than histories of secession or partition.") 
    

```


