---
title: "Climate Colonialism: Gather.Rmd"
author: "Winona Guo"
date: "10/14/2020"
output: html_document
---

These are my datasets:
  
1) Colonial History Dataset
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/5EMETG

2) Measuring the Impacts of Colonialism: A New Data Set for the Countries of Africa and Asia
https://dataverse.harvard.edu/dataset.xhtml?persistentId=doi:10.7910/DVN/UQZFYA

3) Countries Vulnerable to the Climate
https://www.cgdev.org/publication/dataset-vulnerability-climate-change

4) C02 Emissions Per Country
https://github.com/datasets/co2-fossil-by-nation

Planning out my sitemap:

1) The Challenge
2) History of Empire
3) Emissions vs. Effects
4) Model
5) About

Below are resources that were helpful to reference as I made this project:

https://shiny.rstudio.com/reference/shiny/1.5.0/
https://shiny.rstudio.com/tutorial/
https://github.com/wyatthurt/gov-50-class-scripts/blob/main/class-2020-09-22/week_3_A.Rmd
https://rdrr.io/cran/maptools/man/wrld_simpl.html
https://www.davidkane.info/files/final_projects.html
https://github.com/wyatthurt/transboundary-water-conflict/tree/master/shiny_app
https://www.r-graph-gallery.com/

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(shiny)
library(shinythemes)
library(tidyverse)
library(janitor)

# Janitor is used to the clean the names within my data.

library(readr)
library(readxl)
library(leaflet)
library(maptools)
library(rgdal)
library(dplyr)
library(tidymodels)
library(rstanarm)
library(gt)
library(broom.mixed)

```

```{r Loading Data}

colonialism <- read_xls("shiny_app/raw_data/colonialism.xls") %>%
  clean_names() 

climaterisk <- read_xls("shiny_app/raw_data/vulnerability.xls", skip = 2) %>%
  clean_names() %>%
  rename(vulnerable = climate_vulnerability_cv_cdi_adj_for_income_regulation) %>%
  rename(governance = kkm_regulatory_quality_score_2008) %>%
  rename(income = income_per_capita_us_ppp_2008) %>%
  rename(area = area_sq_km) %>%
  mutate(country = str_to_title(country)) %>%
  select(country, vulnerable, income, governance, area, world_sub_region) %>%
  mutate(country = recode(country, 'Korea, Rep.' = 'Korea, Republic of',
                          "Korea, Dem. Rep." = "Korea, Democratic People's Republic of",
                          'Taiwan (China)' = 'Taiwan',
                          'Antigua And Barbuda' = 'Antigua and Barbuda',
                          'Bahamas, The' = 'Bahamas',
                          'Slovak Republic' = 'Slovakia',

                          # Starting here are the ones I am renaming so that it aligns with the wrldsimplmod countries' names. 

                          'Iran, Islamic Rep.' = 'Iran (Islamic Republic of)',
                          'Bosnia And Herzegovina' = 'Bosnia and Herzegovina',
                          'Myanmar' = 'Burma',
                          'Congo, Rep.' = 'Congo',
                          'Congo, Dem. Rep.' = 'Democratic Republic of the Congo',
                          'Egypt, Arab Rep.' = 'Egypt',
                          'Yemen, Rep.' = 'Yemen',
                          'French Guians' = 'French Guiana',
                          'Micronesia, Fed. Sts.' = 'Micronesia, Federated States of',
                          'Gambia, The' = 'Gambia',
                          "Cote D'ivoire" = "Cote d'Ivoire",
                          'Kyrgyz Republic' = 'Kyrgyzstan',
                          'Falkland Islands' = 'Falkland Islands (Malvinas)',
                          "Lao Pdr" = "Lao People's Democratic Republic",
                          'Libya' = 'Libyan Arab Jamahiriya',
                          'Macedonia, Fyr' = 'The former Yugoslav Republic of Macedonia',
                          'Hong Kong Sar, China' = 'Hong Kong',
                          'Isle Of Man' = 'Isle of Man',
                          'Macao Sar, China' = 'Macau',
                          'West Bank And Gaza' = 'Palestine',
                          
                          # I feel somewhat like I am simplifying decades of history into one line of code!
                          
                          'Serbia And Montenegro' = 'Serbia',
                          'Moldova' = 'Republic of Moldova',
                          'Russian Federation' = 'Russia',
                          'St. Kitts And Nevis' = 'Saint Kitts and Nevis',
                          'St. Lucia' = 'Saint Lucia',
                          'Trinidad And Tobago' = 'Trinidad and Tobago',
                          'Sao Tome And Principe' = 'Sao Tome and Principe',
                          'Tanzania' = 'United Republic of Tanzania',
                          'St. Vincent And The Grenadines' = 'Saint Vincent and the Grenadines',
                          'Venezuela, Rb' = 'Venezuela',
                          'Vietnam' = 'Viet Nam',
                          'Virgin Islands (U.s.)' = 'United States Virgin Islands',
                          'Wallis And Futuna' = 'Wallis and Futuna Islands',
                          'Pitcairn' = 'Pitcairn Islands',
                          'St. Pierre And Miquelon' = 'Saint Pierre and Miquelon',
                          'St. Helena' = 'Saint Helena',
                          'Turks And Caicos Islands' = 'Turks and Caicos Islands',
                          'Svalbard And Jan Mayen' = 'Svalbard')) %>%
    mutate(vulnerable = replace(vulnerable, 188, 42.000))

# BIG IMPORTANT NOTE: I replaced Somalia's 100.000 climate risk rate with 42.000 (near Burma's), because it is a single dramatic outlier that changes the later graph. I will be sure to note this via text inside my Shiny app!

# Mapping these data sets (and doing the corresponding recoding work) requires navigating a lot of complex history that I'm having difficulty summarizing into one number. For instance, in the countries data, we have Yemen listed as being decolonized in 1990, but Yemen People’s Republic decolonized in 1967. In the climate data, there is one one 'Yemen Rep.' Which name maps on to which? I ended up choosing to rename Yemen Rep. as Yemen, both to match the later colonization date and also to match wrldsimplmod.

countries <- read_xls("shiny_app/raw_data/countries_files/icowcol.xls") %>%
  mutate(date = str_sub(Indep, 1, 4)) %>%
  mutate(twenty = ifelse(date >= 1900, "Colonized", "Independent")) %>%
  mutate(Name = recode(Name, 'United States of America' = 'United States',
                       'Tunisia (postcolonial)' = 'Tunisia',
                       'St. Kitts and Nevis' = 'Saint Kitts and Nevis',
                       'St. Vincent and the Grenadines' ='Saint Vincent and the Grenadines',
                       'Prussia / Germany' = 'Germany',
                       'Fed. States of Micronesia' = 'Micronesia, Federated States of',
                       'Iran' = 'Iran (Islamic Republic of)',
                       'Brunei' = 'Brunei Darussalam',
                       'Austria-Hungary' = 'Austria',
                       "Laos" = "Lao People's Democratic Republic",
                       'Egypt (poat-colonial)' = 'Egypt',
                       'South Korea' = 'Korea, Republic of',
                       "Korea, Dem. Rep." = "Korea, Democratic People's Republic of",
                       'Moldova' = 'Republic of Moldova',
                       'Serbia / Yugoslavia' = 'Serbia',
                       'Tanzania' = 'United Republic of Tanzania',
                       'Vietnam' = 'Viet Nam'))

emissions <- read_csv("shiny_app/raw_data/emissions.csv", 
                      col_types = cols(
                        Year = col_double(),
                        Country = col_character(),
                        Total = col_double(),
                        `Solid Fuel` = col_double(),
                        `Liquid Fuel` = col_double(),
                        `Gas Fuel` = col_double(),
                        Cement = col_double(),
                        `Gas Flaring` = col_double(),
                        `Per Capita` = col_double(),
                        `Bunker fuels (Not in Total)` = col_double()
                      )) %>%
  group_by(Country) %>%
  mutate(sum = sum(Total)) %>%
  slice(1) %>%
  select(Country, sum) %>%
  mutate(Country = str_to_title(Country)) %>%
  mutate(Country = recode(Country, "United States Of America" = "United States",
                          "China (Mainland)" = "China",
                          "Russian Federation" = "Russia",
                          "Islamic Republic Of Iran" = "Iran (Islamic Republic of)",
                          "Plurinational State Of Bolivia" = "Bolivia",
                          "Myanmar (Formerly Burma)" = "Burma",
                          "United Republic Of Tanzania" = "United Republic of Tanzania",
                          "Libyan Arab Jamahiriyah" = "Libyan Arab Jamahiriya",
                          "French West Africa" = "Western Sahara",
                          "Cote D Ivoire" = "Cote d'Ivoire",
                          "Republic Of Cameroon" = "Cameroon",
                          "Democratic Republic Of The Congo (Formerly Zaire)" = "Democratic Republic of the Congo",
                          "United Republic Of Tanzania" = "United Republic of Tanzania",
                          "Lao People S Democratic Republic" = "Lao People's Democratic Republic",
                          "Republic Of Korea" = "Korea, Republic of", 
                          "Democratic People S Republic Of Korea" = "Korea, Democratic People's Republic of",
                          "Antarctic Fisheries" = "Antarctica",
                          "France (Including Monaco)" = "France",
                          "Italy (Including San Marino)" = "Italy")) %>%
    mutate(sum = ifelse(sum == 102510260, 60000000, sum))

# BIG IMPORTANT NOTE AGAIN: I replaced the U.S. total emissions of 102510260, a dramatic outlier, to be 60,000,000, closer to the second largest # of 47649834 (China). 

# The below joins the colonialism data with climate risk data. 

fit_mod <- inner_join(countries, climaterisk, by = c("Name" = "country")) %>%
 group_by(twenty) %>%
 mutate(avg_risk = mean(vulnerable)) 

```

```{r Tab 1 Plot}

# Count(twenty) yields 131 colonized and 45 independent countries.

bar_graph <- fit_mod %>%
 select(twenty, avg_risk) %>%
 slice(1) %>%
 ggplot(mapping = aes(x = twenty, y = avg_risk)) + 
 geom_col(fill = "darkgreen") +
 labs(title = "Climate Risk Today for Countries Colonized vs. Independent", 
      subtitle = "At The Turn of the 20th Century", 
      x = "Countries' Colonial Status in 1900",
      y = "Average Climate Vulnerability Today",
      caption = "The average climate risk for colonized countries is 4.58, 
      and the average for independent countries is 1.52.") +
 theme_classic()

bar_graph

```

``` {r Tab 2 Plots}

# PLOT 1: COLONIALISM & INCOME

#graph onset and end of colonialism for different countries
#graph the 4 variables change over that period
#graph france vs. uk whether they have different effects on climate risk


#countries on left, geomline line segments on y from onset/end, color coded by motherland

#renacohen.shinyapps.io/afta/
 # https://benalexkeen.com/creating-a-timeline-graphic-using-r-and-ggplot2/

#look at rena's shiny app for modeling results

# PLOT 2: NUMBER OF YEARS COLONIZED (African and Asian Countries)

q1 <- colonialism %>%
  select(country_name, colyears) 

q2 <- climaterisk %>%
  select(country, vulnerable, world_sub_region) %>%
  subset(country != "Somalia") %>%
  subset(country != "Burundi")

# I removed these two outliers because they were so extreme that the rest of the graph was difficult to decipher.

q3 <- inner_join(q1, q2, by = c("country_name" = "country")) 

colonize <- q3 %>%
  ggplot(aes(x = colyears, y = vulnerable, color = world_sub_region)) +
  geom_point(alpha = 0.7) +
  labs(title = "Countries by Years Colonized and Vulnerability to Climate Risk",
       x = "Years Colonized",
       y = "Vulnerability to Climate Risk",
       caption = "This graph indicates there is not a significant correlation
       overall between the years a country has been colonized, 
       and its current vulnerability to climate risk.") +
  theme_bw() + 
  scale_color_discrete("World Region")

# PLOT: TYPE OF INDEPENCE (x) ON CLIMATE RISK (y)
      
    types <- fit_mod %>%
      ggplot(aes(x = Type, y = vulnerable)) +
      geom_col(fill = "orange") +
      labs(title = "The Effect of Different Types of Colonialism on Climate Risk",
           y = "Climate Risk",
           subtitle = "Key: 1 = Formation, 2 = Decolonization, 3 = Secession, 4 = Partition",
           caption = "This graph shows that decolonialization and occuption by a foreign power 
           has a significantly greater effect on risk than histories of secession or partition.") 
    
    types
    

```
```{r Tab 3 Leaflet}

# Modifying Polygons: world countries from the maptools package.

data(wrld_simpl)

wrldsimplmod <- wrld_simpl

# Merging with my data on climate risk. Unfortunately some countries are left with NA's because there is no climate risk data. An example: the South Georgia South Sandwich Islands!

wrldsimplmod@data <-
  
   left_join(wrld_simpl@data, climaterisk, by = c("NAME" = "country"))

saveRDS(wrldsimplmod, "shiny_app/joined.rds")

# Making for the second graph.

secondmod <- wrld_simpl

secondmod@data <-
  
   left_join(wrld_simpl@data, emissions, by = c("NAME" = "Country"))

saveRDS(secondmod, "shiny_app/joined2.rds")

```

```{r Tab 4 Model}

# Is there a predictive relationship between CO2 emissions and climate risk?

# Trying to produce total emissions for each country between 1751 and 2014. 

model <- inner_join(emissions, climaterisk, by = c("Country" = "country"))

model %>%
  ggplot(aes(x = sum, y = vulnerable)) +
  geom_smooth(method = "lm") +
  scale_x_log10() +
  scale_y_log10()

# possibly do checkboxes to give people choices
#show geom point not smooth to show variability
#specify these r countries that have been colonized (look back at 'model' what is missing). i see little correlation because...

# not 1 x 10^7 but log 1 x 10^7

# I am also controlling for three variables from each country here: income per capita, area, and governance quality. This data was included inside the climate risk dataset.

reg <- stan_glm(formula = vulnerable ~ sum + income + area + governance,
                data = model, 
                refresh = 0, 
                seed = 8) %>%
  tidy()

# create model with 0 and 1 as colonized vs indepdent. run that on climate risk - start here! if i add twenty column it will convert for me!!!!!!!!!!

# include model tibble via renderDataTable in shinyapp (see wyatt's first page!)


```

* Write the math for this model. I am considering risk as a function of emissions. How much should/do emissions affect risk?

$$ reg_{risk} = \beta_0 + \beta_1x_{emissions, i} + \epsilon $$
















